---
title: "Mini Data-Analysis Deliverable 1"
output: github_document
---

# Welcome to your (maybe) first-ever data analysis project!

And hopefully the first of many. Let's get started:

1.  Install the [`datateachr`](https://github.com/UBC-MDS/datateachr) package by typing the following into your **R terminal**:

<!-- -->

```         
install.packages("devtools")
devtools::install_github("UBC-MDS/datateachr")
```

2.  Load the packages below.

```{r}
library(datateachr)
library(tidyverse)
# To facilitate parsing
library(stringr)
# Nice colors
library(RColorBrewer)
# Scale axis
library(scales)

```

3.  Make a repository in the <https://github.com/stat545ubc-2023> Organization. You can do this by following the steps found on canvas in the entry called [MDA: Create a repository](https://canvas.ubc.ca/courses/126199/pages/mda-create-a-repository). One completed, your repository should automatically be listed as part of the stat545ubc-2023 Organization.

# Instructions

## For Both Milestones

-   Each milestone has explicit tasks. Tasks that are more challenging will often be allocated more points.

-   Each milestone will be also graded for reproducibility, cleanliness, and coherence of the overall Github submission.

-   While the two milestones will be submitted as independent deliverables, the analysis itself is a continuum - think of it as two chapters to a story. Each chapter, or in this case, portion of your analysis, should be easily followed through by someone unfamiliar with the content. [Here](https://swcarpentry.github.io/r-novice-inflammation/06-best-practices-R/) is a good resource for what constitutes "good code". Learning good coding practices early in your career will save you hassle later on!

-   The milestones will be equally weighted.

## For Milestone 1

**To complete this milestone**, edit [this very `.Rmd` file](https://raw.githubusercontent.com/UBC-STAT/stat545.stat.ubc.ca/master/content/mini-project/mini-project-1.Rmd) directly. Fill in the sections that are tagged with `<!--- start your work below --->`.

**To submit this milestone**, make sure to knit this `.Rmd` file to an `.md` file by changing the YAML output settings from `output: html_document` to `output: github_document`. Commit and push all of your work to the mini-analysis GitHub repository you made earlier, and tag a release on GitHub. Then, submit a link to your tagged release on canvas.

**Points**: This milestone is worth 36 points: 30 for your analysis, and 6 for overall reproducibility, cleanliness, and coherence of the Github submission.

# Learning Objectives

By the end of this milestone, you should:

-   Become familiar with your dataset of choosing
-   Select 4 questions that you would like to answer with your data
-   Generate a reproducible and clear report using R Markdown
-   Become familiar with manipulating and summarizing your data in tibbles using `dplyr`, with a research question in mind.

# Task 1: Choose your favorite dataset

The `datateachr` package by Hayley Boyce and Jordan Bourak currently composed of 7 semi-tidy datasets for educational purposes. Here is a brief description of each dataset:

-   *apt_buildings*: Acquired courtesy of The City of Toronto's Open Data Portal. It currently has 3455 rows and 37 columns.

-   *building_permits*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 20680 rows and 14 columns.

-   *cancer_sample*: Acquired courtesy of UCI Machine Learning Repository. It currently has 569 rows and 32 columns.

-   *flow_sample*: Acquired courtesy of The Government of Canada's Historical Hydrometric Database. It currently has 218 rows and 7 columns.

-   *parking_meters*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 10032 rows and 22 columns.

-   *steam_games*: Acquired courtesy of Kaggle. It currently has 40833 rows and 21 columns.

-   *vancouver_trees*: Acquired courtesy of The City of Vancouver's Open Data Portal. It currently has 146611 rows and 20 columns.

**Things to keep in mind**

-   We hope that this project will serve as practice for carrying our your own *independent* data analysis. Remember to comment your code, be explicit about what you are doing, and write notes in this markdown document when you feel that context is required. As you advance in the project, prompts and hints to do this will be diminished - it'll be up to you!

-   Before choosing a dataset, you should always keep in mind **your goal**, or in other ways, *what you wish to achieve with this data*. This mini data-analysis project focuses on *data wrangling*, *tidying*, and *visualization*. In short, it's a way for you to get your feet wet with exploring data on your own.

And that is exactly the first thing that you will do!

1.1 **(1 point)** Out of the 7 datasets available in the `datateachr` package, choose **4** that appeal to you based on their description. Write your choices below:

**Note**: We encourage you to use the ones in the `datateachr` package, but if you have a dataset that you'd really like to use, you can include it here. But, please check with a member of the teaching team to see whether the dataset is of appropriate complexity. Also, include a **brief** description of the dataset here to help the teaching team understand your data.

<!-------------------------- Start your work below ---------------------------->

1: *steam_games*\
2: *apt_buildings*\
3: *flow_sample*\
4: *vancouver_trees*

<!----------------------------------------------------------------------------->

1.2 **(6 points)** One way to narrowing down your selection is to *explore* the datasets. Use your knowledge of dplyr to find out at least *3* attributes about each of these datasets (an attribute is something such as number of rows, variables, class type...). The goal here is to have an idea of *what the data looks like*.

*Hint:* This is one of those times when you should think about the cleanliness of your analysis. I added a single code chunk for you below, but do you want to use more than one? Would you like to write more comments outside of the code chunk?

<!-------------------------- Start your work below ---------------------------->

```{r steam_games overview}
# See the name of every column in the dataset
steam_games %>%
  colnames()

# Visualize at a glance the range of values
steam_games %>%
  summary() %>%
  knitr::kable( )

# Observe the information in the first couple of rows
# Removed game description, recommended and minimum requirements for a better table output
steam_games %>%
  select(!game_description) %>%
  select(!minimum_requirements) %>%
  select(!recommended_requirements) %>%
  head(5) %>%
  knitr::kable()
```

```{r apt_buildings overview}
# See the name of every column in the dataset
apt_buildings %>%
  colnames()

# Visualize at a glance the range of values
apt_buildings %>%
  summary() %>%
  knitr::kable()

# Observe the information in the first couple of rows
apt_buildings %>%
  head(5) %>%
  knitr::kable()
```

```{r flow_sample}
# See the name of every column in the dataset
flow_sample %>%
  colnames()

# Visualize at a glance the range of values
flow_sample %>%
  summary() %>%
  knitr::kable()

# Observe the information in the first couple of rows
flow_sample %>%
  head(5) %>%
  knitr::kable()
```

```{r vancouver_trees overview}
# See the name of every column in the dataset
vancouver_trees %>%
  colnames()

# Visualize at a glance the range of values
vancouver_trees %>%
  summary() %>%
  knitr::kable()

# Observe the information in the first couple of rows
vancouver_trees %>%
  head(5) %>%
  knitr::kable()
```

<!----------------------------------------------------------------------------->

1.3 **(1 point)** Now that you've explored the 4 datasets that you were initially most interested in, let's narrow it down to 1. What lead you to choose this one? Briefly explain your choice below.

<!-------------------------- Start your work below ---------------------------->

I'm keeping my first choice of *steam_games*, It would be a nice change to work with something unrelated to biology. I think the main challenge when analyzing this dataset is the parsing of the information in each column. However, there are fields with a very clear structure which will make the task easier.

<!----------------------------------------------------------------------------->

1.4 **(2 points)** Time for a final decision! Going back to the beginning, it's important to have an *end goal* in mind. For example, if I had chosen the `titanic` dataset for my project, I might've wanted to explore the relationship between survival and other variables. Try to think of 1 research question that you would want to answer with your dataset. Note it down below.

<!-------------------------- Start your work below ---------------------------->

I would like to explore if there is any association relationship between game ratings and variables such as:

-   game_details: If the game online, multi-player or single-player

-   genre

-   developer

<!----------------------------------------------------------------------------->

# Important note

Read Tasks 2 and 3 *fully* before starting to complete either of them. Probably also a good point to grab a coffee to get ready for the fun part!

This project is semi-guided, but meant to be *independent*. For this reason, you will complete tasks 2 and 3 below (under the **START HERE** mark) as if you were writing your own exploratory data analysis report, and this guidance never existed! Feel free to add a brief introduction section to your project, format the document with markdown syntax as you deem appropriate, and structure the analysis as you deem appropriate. If you feel lost, you can find a sample data analysis [here](https://www.kaggle.com/headsortails/tidy-titarnic) to have a better idea. However, bear in mind that it is **just an example** and you will not be required to have that level of complexity in your project.

# Task 2: Exploring your dataset

If we rewind and go back to the learning objectives, you'll see that by the end of this deliverable, you should have formulated *4* research questions about your data that you may want to answer during your project. However, it may be handy to do some more exploration on your dataset of choice before creating these questions - by looking at the data, you may get more ideas. **Before you start this task, read all instructions carefully until you reach START HERE under Task 3**.

2.1 **(12 points)** Complete *4 out of the following 8 exercises* to dive deeper into your data. All datasets are different and therefore, not all of these tasks may make sense for your data - which is why you should only answer *4*.

Make sure that you're using dplyr and ggplot2 rather than base R for this task. Outside of this project, you may find that you prefer using base R functions for certain tasks, and that's just fine! But part of this project is for you to practice the tools we learned in class, which is dplyr and ggplot2.

1.  Plot the distribution of a numeric variable.
2.  **Create a new variable based on other variables in your data (only if it makes sense)**
3.  **Investigate how many missing values there are per variable. Can you find a way to plot this?**
4.  **Explore the relationship between 2 variables in a plot.**
5.  Filter observations in your data according to your own criteria. Think of what you'd like to explore - again, if this was the `titanic` dataset, I may want to narrow my search down to passengers born in a particular year...
6.  Use a boxplot to look at the frequency of different observations within a single variable. You can do this for more than one variable if you wish!
7.  **Make a new tibble with a subset of your data, with variables and observations that you are interested in exploring.**
8.  Use a density plot to explore any of your variables (that are suitable for this type of plot).

2.2 **(4 points)** For each of the 4 exercises that you complete, provide a *brief explanation* of why you chose that exercise in relation to your data (in other words, why does it make sense to do that?), and sufficient comments for a reader to understand your reasoning and code.

<!-------------------------- Start your work below ---------------------------->

### Data pre-processing

The first goal is to observe the variability of the review score. If all reviews are too homogeneous, it would be difficult to associate this variable with any other. For this reason, I will create several new columns as follow:

-   Categorical rating: Mixed, Good, etc

-   Fraction of positive reviews: Convert the percentage to a fraction for easier handling

-   Total number of reviews: Extract the total number of reviews

-   Number of positive reviews: Multiply the fraction by the total number of reviews

I will focus on "all_reviews" since it will be a more fair comparison for games that have not had a lot of recent reviews. By doing this, I acknowledge that some games rating can change overtime. I don't know how frequent this trend is, but this selection seemed like the "lesser evil".

```{r}
reviews_data <- steam_games %>%
  select(all_reviews) %>%
  # Extract possible category choices
  mutate(categorical = str_replace(all_reviews, ",\\(.+", "")) %>%
  # Extract percentage of positive reviews and convert to fraction
  mutate(pos_fraction = str_extract(all_reviews, "[0-9]+%")) %>%
  mutate(pos_fraction = as.numeric(str_replace(pos_fraction, "%", ""))/100) %>%
  # Extract the total number of reviews
  mutate(total_reviews = str_extract(all_reviews, "of the .+ user")) %>%
  mutate(total_reviews = str_replace(total_reviews, "of the ", "")) %>%
  mutate(total_reviews = str_replace(total_reviews, " user", "")) %>%
  mutate(total_reviews = as.numeric(str_replace(total_reviews, ",", "")))

# Visualize what other stuff got into the categorical variable
unique(reviews_data$categorical)
# Select the relevant categories
valid_categories <- unique(reviews_data$categorical)[c(7,5,3,1,2,19,10,20,18)]
# Final format
reviews_data <- reviews_data %>%
  # Extract low user-reviews and assign it to the corresponding variable
  mutate(total_reviews = ifelse(str_detect(categorical, "user reviews"),
                                str_extract(categorical, "^[0-9]+"),
                                total_reviews)) %>%
  # Conserve valid categories
  mutate(categorical = factor(categorical, levels = valid_categories)) %>%
  mutate(total_reviews = as.numeric(total_reviews)) %>%
  # Calculate number of positive reviews, rounded up
  mutate(pos_reviews = ceiling(pos_fraction * total_reviews)) %>%
  select(!all_reviews)

reviews_data %>%
  head() %>%
  knitr::kable()

```

### Data visualization

I will plot the frequency of the categorical rating across all games:

```{r}

category_colors <- brewer.pal(9, "RdYlGn")[seq(9,1)]

reviews_data %>%
  ggplot(aes(x = categorical, fill = categorical)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values = category_colors)+
  theme_classic() +
  labs(x = "")

```

Since there is a lot of NA values, I repeated the same visualization but now removing the NAs

```{r}
reviews_data %>%
  na.omit() %>%
  ggplot(aes(x = categorical, fill = categorical)) +
  geom_bar() +
  coord_flip() +
  scale_fill_manual(values = category_colors)+
  theme_classic() +
  labs(x = "")
```

I also visualized the distribution of the fraction of positive reviews.

```{r}
reviews_data %>%
  ggplot(aes(x = pos_fraction)) +
  geom_histogram(binwidth = 0.05)
```

Finally I observed the relationship between number of reviews and fraction of positive reviews. Since having few reviews in general can artificially increase the effect of negative reviews.

```{r}
reviews_data %>%
  na.omit() %>%
  ggplot(aes(x = pos_fraction, y = total_reviews + 1)) +
  geom_point() +
  scale_y_continuous(trans = "log10", labels = comma) +
  labs(x = "Fraction of positive reviews", y = "log10( Total number of reviews )")

```

There seems to be a slight bias of bad reviews and games with less than 100 reviews.

```{r}
reviews_data %>%
  na.omit() %>%
  filter(total_reviews > 100) %>%
  ggplot(aes(x = pos_fraction, y = total_reviews + 1)) +
  geom_point() +
  scale_y_continuous(trans = "log10", labels = comma) +
  labs(x = "Fraction of positive reviews", y = "log10( Total number of reviews )")
```

There still seems to be a slight bias of games with very bad reviews and less than a 1,000 reviews.

One idea to start exploring the relationship of the positive review fraction with the original game price

```{r}
reviews_data %>%
  mutate(og_price = round(as.numeric(steam_games$original_price))) %>%
  filter(total_reviews > 100) %>%
  mutate(og_price = ifelse(og_price > 100, 100, og_price)) %>%
  na.omit() %>%
  ggplot(aes(x = og_price)) +
  geom_histogram(binwidth = 5) +
  labs(x = "Launch price - truncated to 100")
```

```{r}
reviews_data %>%
  mutate(og_price = round(as.numeric(steam_games$original_price))) %>%
  filter(total_reviews > 100) %>%
  na.omit() %>%
  ggplot(aes(x = pos_fraction, y = og_price  + 1)) +
  geom_point() +
  scale_y_continuous(trans = "log10", labels = comma) +
  labs(x = "Fraction of positive reviews", y = "log10( Launch price + 1 )")

```

There doesn't seem to be a strong association between launch price and the fraction of positive reviews, but further testing is needed to confirm.

<!----------------------------------------------------------------------------->

# Task 3: Choose research questions

**(4 points)** So far, you have chosen a dataset and gotten familiar with it through exploring the data. You have also brainstormed one research question that interested you (Task 1.4). Now it's time to pick 4 research questions that you would like to explore in Milestone 2! Write the 4 questions and any additional comments below.

<!--- *****START HERE***** --->

The initial exploration does indicate any strong associations between fraction of positive reviews and either the launch price or the number of reviews. The abundance of positive reviews in general could be an indicator that people is more willing to write positive reviews than negative comments. However, I think it still worth to evaluate if positive reviews are associated with other variables in particular:

-   Player count: single-player vs multi-player

-   Game genre

-   Game publisher / game developer

-   Length of comment

I think it would be interesting to test if there is any statistical association between those variables and fraction of positive comments.

<!----------------------------->

# Overall reproducibility/Cleanliness/Coherence Checklist

## Coherence (0.5 points)

The document should read sensibly from top to bottom, with no major continuity errors. An example of a major continuity error is having a data set listed for Task 3 that is not part of one of the data sets listed in Task 1.

## Error-free code (3 points)

For full marks, all code in the document should run without error. 1 point deduction if most code runs without error, and 2 points deduction if more than 50% of the code throws an error.

## Main README (1 point)

There should be a file named `README.md` at the top level of your repository. Its contents should automatically appear when you visit the repository on GitHub.

Minimum contents of the README file:

-   In a sentence or two, explains what this repository is, so that future-you or someone else stumbling on your repository can be oriented to the repository.
-   In a sentence or two (or more??), briefly explains how to engage with the repository. You can assume the person reading knows the material from STAT 545A. Basically, if a visitor to your repository wants to explore your project, what should they know?

Once you get in the habit of making README files, and seeing more README files in other projects, you'll wonder how you ever got by without them! They are tremendously helpful.

## Output (1 point)

All output is readable, recent and relevant:

-   All Rmd files have been `knit`ted to their output md files.
-   All knitted md files are viewable without errors on Github. Examples of errors: Missing plots, "Sorry about that, but we can't show files that are this big right now" messages, error messages from broken R code
-   All of these output files are up-to-date -- that is, they haven't fallen behind after the source (Rmd) files have been updated.
-   There should be no relic output files. For example, if you were knitting an Rmd to html, but then changed the output to be only a markdown file, then the html file is a relic and should be deleted.

(0.5 point deduction if any of the above criteria are not met. 1 point deduction if most or all of the above criteria are not met.)

Our recommendation: right before submission, delete all output files, and re-knit each milestone's Rmd file, so that everything is up to date and relevant. Then, after your final commit and push to Github, CHECK on Github to make sure that everything looks the way you intended!

## Tagged release (0.5 points)

You've tagged a release for Milestone 1.

### Attribution

Thanks to Icíar Fernández Boyano for mostly putting this together, and Vincenzo Coia for launching.
